---
title: "Week 01 - Intro to Environmental Data Science in R"
author: "Daniel Rode"
date: "1/20/2023"
output: html_document
---


```{r}
source("setup.R")
```


# Exercise 6.3.1 - 1
```{r}
# create myList list
myList <- list("apple", 1993, FALSE, penguins)
list(myList, list("more stuff here", list("and more")))
names(myList) <- c("fruit", "year", "logic", "data")


# error:
myList["Fruit"]
# fix:
myList[["fruit"]]

# error:
penguins$flipper_lenght_mm
# fix:
penguins$flipper_length_mm

# error: 
penguins[island=='Dream',]
# fix:
penguins[penguins$island=='Dream',]
```


# Exercise 6.3.1 - 2
There are 3 species of penguins in the `penguins` dataset, as determined by:

```{r}
length(unique(penguins$species))
```


# Exercise 6.3.1 - 3
Pruning `penguins` to species, island, and flipper_length_mm columns:

```{r}
penguins %>%
  select(species, island, flipper_length_mm) %>%
  filter(island=='Dream')
```


# Exercise 6.3.1 - 4
Mean flipper length of the *Adelie* species is 190 mm, as determined by:

```{r}
penguins[penguins$species=='Adelie',]$flipper_length_mm %>% 
  mean(na.rm=TRUE)
```


# Exercise 6.3.2 - 1
> Reorder the variables in penguins so that year is the first column followed by the rest (Hint: look into the use of everything()).

```{r}
year_col = select(penguins, year)
remaining_cols = select(penguins, -year)
penguins = mutate(year_col, remaining_cols)

# probably could have just used relocate()
```


# Exercise 6.3.2 - 2
> Create a new column called ‘size_group’ where individuals with body mass greater than the overall average are called ‘large’ and those smaller are called ‘small’.

```{r}
avg_mass = mean(penguins$body_mass_g, na.rm=TRUE)

# mutate based on conditional statements
penguins = penguins %>% 
  mutate(
    size_group = if_else(
      body_mass_g < avg_mass, 
      "small", "large"
  )
)
```


# Exercise 6.3.2 - 3
> Find out which year for each species were individuals on average the largest according to body mass.

```{r}
years = unique(penguins$year)
for(i in 1:length(years)) {
  yr = years[i]
  print(yr)
}
# the code above is not used to solve a problem, but it will remind me how to iterate/loop
```

Individuals of the *Adelie* species were largest (on average) during 2008 (within the span of the given dataset). For *Chinstrap*, the year was 2008, and for *Gentoo*, the year was 2009. See:
```{r}
penguins %>%
  group_by(year, species) %>%
  summarise(mean_mass = mean(body_mass_g, na.rm=TRUE))
```


# Exercise 6.3.2 - 4
> You want to filter data for years that are not in a vector of given years, but this code doesn’t work. Tweak it so that it does. (Yes, you could just filter year to equal 2007 in this case but there is a trouble-shooting lessons here).

```{r}
penguins %>% 
  filter(!year %in% c(2008, 2009))
```


# Exercise 7.1 - 1

```{r}
library(ggplot2)
library(showtext)

showtext_auto()  # Fixes font issues on Linux

# the harder way:
# penguins %>%
#   group_by(species) %>%
#   rename_at("species", ~"Species") %>%  # Rename column on the fly
#   summarise(
#     "Mean Flipper Length (mm)" = mean(
#       flipper_length_mm,
#       na.rm=TRUE
#       )
#     ) %>% 
#   ggplot() +
#   geom_bar(mapping = aes(x = `Mean Flipper Length (mm)`, fill = Species))

ggplot(penguins) +
  geom_bar(
    stat = "summary",
    fun = "mean",
    mapping = aes(
      x = species,
      y = flipper_length_mm),
    ) +
  labs(
    title = "Penguins",
    ) +
  xlab("Species") +
  ylab("Mean Flipper Length (mm)")
```


# Exercise 7.1 - 2

```{r}
penguins %>%
  filter(island=='Dream') %>%
  ggplot() +
    geom_point(mapping = aes(
      x = bill_length_mm, 
      y = bill_depth_mm, 
      color = species))
```
