---
title: "Week 01 - Intro to Spatial Data in R"
author: "Daniel Rode"
date: "1/23/2023"
output: html_document
---


```{r}
source("setup.R")
```


# Exercise 1
> Filter out the counties data set to only include Larimer, Denver, and Pueblo counties.

```{r}
counties = counties(state = "CO")
filter(counties, NAME %in% c("Larimer", "Denver", "Pueblo"))
```


# Exercise 2
> Make a map of the counties data colored by county area. Make a second map of counties colored by their total area of water.

```{r}
counties = mutate(counties, AREA = ALAND+AWATER)

qtm(counties, fill = "AREA")

qtm(counties, fill = "AWATER")
```


# Exercise 3
> Make a barplot comparing the elevation of your 3 points in the Poudre Canyon (note: explore the extract() function in the terra package).

```{r}
# Elevation units are in meters
elevation = get_elev_raster(counties, z = 7)

# Isolate and name several locations
poudre_points = data.frame(
  name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
  long = c(-105.35634, -105.58159, -105.85563),
  lat = c(40.68752, 40.69687, 40.57960)
)
poudre_points_sf = st_as_sf(poudre_points, coords = c("long", "lat"), crs = 4326)  # Convert to sf object

# Extract elevation for locations isolated above, and save as column to sf object
poudre_points_sf$elevation_m = extract(elevation, poudre_points_sf)

# Plot elevations
ggplot(
  poudre_points_sf,
  mapping = aes(x = name, y = elevation_m),
  ) +
  geom_bar(stat = "identity") +
  labs(title = "Elevation of 3 Points in Poudre Canyon") +
  xlab("Place") +
  ylab("Elevation (m)")
```


# Exercise 4
> Why are there 4 features in our Poudre Canyon Highway variable instead of 1?

```{r}
roads = roads(state = "CO", county = "Larimer")
poudre_hwy = roads %>% 
  filter(FULLNAME == "Poudre Canyon Hwy")

qtm(poudre_hwy[1,])
qtm(poudre_hwy[2,])
qtm(poudre_hwy[3,])
qtm(poudre_hwy[4,])
qtm(poudre_hwy)
```

The Poudre Canyon Highway object is composed of 4 features because each of the four lines is a sub-feature of the overall highway (four line-segments that compose the overall line). In reality, I do not know (that is just a guess that is assumed by mapping each feature individually, as done above).
