---
title: "Main"
author: "Daniel Rode"
date: "1/20/2023"
output: html_document
---


```{r}
source("setup.R")
```


# Exercises 6.3.1

## Exercise 1
```{r}
# create myList list
myList <- list("apple", 1993, FALSE, penguins)
list(myList, list("more stuff here", list("and more")))
names(myList) <- c("fruit", "year", "logic", "data")


# error:
myList["Fruit"]
# fix:
myList[["fruit"]]

# error:
penguins$flipper_lenght_mm
# fix:
penguins$flipper_length_mm

# error: 
penguins[island=='Dream',]
# fix:
penguins[penguins$island=='Dream',]
```



## Exercise 2
There are 3 species of penguins in the `penguins` dataset, as determined by:

```{r}
length(unique(penguins$species))
```


## Exercise 3
Pruning `penguins` to species, island, and flipper_length_mm columns:

```{r}
penguins %>%
  select(species, island, flipper_length_mm) %>%
  filter(island=='Dream')
```


## Exercise 4
Mean flipper length of the *Adelie* species is 190 mm, as determined by:

```{r}
penguins[penguins$species=='Adelie',]$flipper_length_mm %>% 
  mean(na.rm=TRUE)
```


# Exercises 6.3.2

## Exercise 1
> Reorder the variables in penguins so that year is the first column followed by the rest (Hint: look into the use of everything()).

```{r}
year_col = select(penguins, year)
remaining_cols = select(penguins, -year)
penguins = mutate(year_col, remaining_cols)

# probably could have just used relocate()
```


## Exercise 2
> Create a new column called ‘size_group’ where individuals with body mass greater than the overall average are called ‘large’ and those smaller are called ‘small’.

```{r}
avg_mass = mean(penguins$body_mass_g, na.rm=TRUE)

# mutate based on conditional statements
penguins = penguins %>% 
  mutate(
    size_group = if_else(
      body_mass_g < avg_mass, 
      "small", "large"
  )
)
```


## Exercise 3
> Find out which year for each species were individuals on average the largest according to body mass.

```{r}
years = unique(penguins$year)
for(i in 1:length(years)) {
  yr = years[i]
  print(yr)
}
# the code above is not used to solve a problem, but it will remind me how to iterate/loop
```

Individuals of the *Adelie* species were largest (on average) during 2008 (within the span of the given dataset). For *Chinstrap*, the year was 2008, and for *Gentoo*, the year was 2009. See:
```{r}
penguins %>%
  group_by(year, species) %>%
  summarise(mean_mass = mean(body_mass_g, na.rm=TRUE))
```


## Exercise 4
> You want to filter data for years that are not in a vector of given years, but this code doesn’t work. Tweak it so that it does. (Yes, you could just filter year to equal 2007 in this case but there is a trouble-shooting lessons here).

```{r}
penguins %>% 
  filter(!year %in% c(2008, 2009))
```
